---
title: "DM_project"
author: "Bai Xue"
date: "12/8/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Librarys
```{r}
library(ggplot2)
library(plyr)
library(ISLR)
library(MASS)
library(knitr)
library(gam)
library(readr)
library(glmnet)
library(rmarkdown)
```

## Read in data and read convert values to 0/1/NA
```{r}
#linear regression
#elementary <- read_csv("Documents/Carnegie Mellon/Fall 2017/Data Mining/Data Mining Project/elementary.csv")
#elementary <- read_csv("/Users/xb/Desktop/Data Mining project/elementary.csv")
elementary <- read_csv("/Users/leonshi/Desktop/Data Mining/Final Project/elementary.csv")
middle <- read_csv("/Users/leonshi/Desktop/Data Mining/Final Project/middle.csv")
high <- read_csv("/Users/leonshi/Desktop/Data Mining/Final Project/high.csv")

elementary.reg <- transform(elementary, 
                  API = mapvalues(API, c("Low","High","?"), 
                                     c(0, 1, NA)))
```


#Change ? to NA
```{r}
#Change ? to NA
elementary.reg[elementary.reg == "?"] <- NA
#middle.reg[middle.reg == "?"] <- NA

#This removes rows where API has null values
elementary.reg <- elementary.reg[-which(is.na(elementary.reg$API)), ]
```


## ACS_CORE
## Change to numeric
```{r}
#Change to numeric 
elementary.reg$ACS_CORE <- as.numeric(elementary.reg$ACS_CORE)
```

#Find the median
```{r}
#Find the median
median.ASC<- median(elementary.reg$ACS_CORE, na.rm = TRUE)
max.ASC <- max(elementary.reg$ACS_CORE,na.rm = TRUE)
elementary.reg$ACS_CORE <- cut(elementary.reg$ACS_CORE, c(1, median.ASC + 1, max.ASC),labels = FALSE)
elementary.reg$ACS_CORE[is.na(elementary.reg$ACS_CORE)] <- 3
```

#change ACS_CORE to categorical
```{r}
#change to categorical
elementary.reg$ACS_CORE <- factor(elementary.reg$ACS_CORE)

```

# Change Charter and Year Round to categorical variable
```{r}
# Change Charter
elementary.reg$CHARTER[is.na(elementary.reg$CHARTER)] <- 0
elementary.reg$CHARTER[elementary.reg$CHARTER == "Y"] <- 1
elementary.reg$CHARTER[elementary.reg$CHARTER == "D"] <- 2

elementary.reg$YR_RND[is.na(elementary.reg$YR_RND)] <- 2
elementary.reg$YR_RND[elementary.reg$YR_RND == "Yes"] <- 0
elementary.reg$YR_RND[elementary.reg$YR_RND == "No"] <- 1

#change to categorical
elementary.reg$CHARTER <- factor(elementary.reg$CHARTER)
elementary.reg$YR_RND <- factor(elementary.reg$YR_RND)
#test
is.factor(elementary.reg$CHARTER)
is.factor(elementary.reg$YR_RND)

```

#Round and change average education to categorical variable
```{r}
#Change null values to 6
elementary.reg$AVG_ED[is.na(elementary.reg$AVG_ED)] <- 6
elementary.reg$AVG_ED <- as.numeric(elementary.reg$AVG_ED)
elementary.reg$AVG_ED <- round(elementary.reg$AVG_ED, digits = 0)

elementary.reg$AVG_ED <- as.factor(elementary.reg$AVG_ED)
is.factor(elementary.reg$AVG_ED)
```


#Converting the Columns 
```{r}
#Converting the Columns 
elementary.reg$API <- factor(elementary.reg$API)
elementary.reg$VALID <- as.numeric(elementary.reg$VALID)
elementary.reg$NUMTEACH <- as.numeric(elementary.reg$NUMTEACH)
elementary.reg$FULL_PCT <- as.numeric(elementary.reg$FULL_PCT)
elementary.reg$EMER_PCT <- as.numeric(elementary.reg$EMER_PCT)
elementary.reg$WVR_PCT <- as.numeric(elementary.reg$WVR_PCT)
elementary.reg$YRS_TEACH <- as.numeric(elementary.reg$YRS_TEACH)
elementary.reg$YRONE_TCH <- as.numeric(elementary.reg$YRONE_TCH)
elementary.reg$YRTWO_TCH <- as.numeric(elementary.reg$YRTWO_TCH)
```

#Logistic regression 

```{r}
#Logistic regression 
elementary.log <- glm(API ~., family = binomial, data = elementary.reg) 
elemLog.summary <- summary(elementary.log)
kable(elemLog.summary$coefficients)
```
#Fitting a lasso on our model
```{r}
#removing all rows where API is null
elementary.reg2 = elementary.reg
#elementary.reg2$API <- NULL

x <- model.matrix(API~.,elementary.reg2)[,-1]
y <- as.matrix(data.frame(elementary.reg$API))
#Need to change y to numeric to run cross validation
y <- as.numeric(y)

# Split data into test and train
set.seed(1)
elem.train <- sample(1:nrow(x), nrow(x)/2)
elem.test <- (-elem.train)
y.test <- y[test]

# If we want to specify the number of lambda values - not sure we actually need this
# Predefined grid of lambda values: 
# grid=10^seq(10,-2, length =100)

#Creating our lasso function
elem.glm <- glmnet(x[elem.train,],y[elem.train],family = "binomial",alpha=1) 
elem.glm
#Plot the training lasso
plot(elem.glm)

#Fit the model on our test data
set.seed(1)
elem.out=cv.glmnet(x[elem.train,],y[elem.train],alpha=1)
plot(elem.out)
```

#Determining coeffcients and significant predictors using lasso
```{r}
#Figuring out the minimum lambda value and predicting CV error on test
elem.bestlam=elem.out$lambda.min
elem.1se=elem.out$lambda.1se
elem.lasso.pred=predict(elem.glm,s=elem.bestlam,newx=x[test,])
print("Minimum lambda for Elementary")
elem.bestlam
print("1SE lambda for Elementary")
elem.1se
print("CV Error for Elementary")
mean((elem.lasso.pred-y.test)^2)

#Determine which variables are non-zero at the 1se value of lambda
out=glmnet(x,y,alpha=1,family="binomial")
elem.lasso.coef=predict(out,type="coefficients",s=elem.1se)
elem.lasso.coef
elem.lasso.coef[elem.lasso.coef!=0]
```

#Forward Selection

```{r}

```

#To do......
#Convert year round to categorical-DONE LS
#Round Education and turn to categorical (current null = 6)
#plot(elementary.log)
#Check conditional probabilities with 
#Classifiers
#Logistic Lasso (Which ones are selected that are different than zero): Convert model to matrix model. Model.matrix. 
#Forward Selection.
#Clustering

## Including Plots

```{r}

```


You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
